# Data Visualization

```{r}
#| output: false
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

## 1.2.5 Exercises

1. How many rows are in penguins? How many columns?

```{r}
print(nrow(penguins))
print(ncol(penguins))
```

2. What does the bill_depth_mm variable in the penguins data frame describe?
   Read the help for ?penguins to find out.

It measures the depth in mm of the bills of the penguins. The depth is the
height of the bill, as opposed to the length. Bill width is not measured.

3. Make a scatterplot of bill_depth_mm vs. bill_length_mm. That is, make a
   scatterplot with bill_depth_mm on the y-axis and bill_length_mm on the
   x-axis. Describe the relationship between these two variables.

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = species)
) +
  geom_point(mapping = aes(shape = species), na.rm = TRUE) +
  geom_smooth(method = "lm", na.rm = TRUE) +
  labs(
    title = "Penguin Bill Length vs. Depth (mm)",
    subtitle = "Segmented by species: Adelie, Chinstrap, & Gentoo",
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)",
    color = "Species", shape = "Species"
  )
```

4. What happens if you make a scatterplot of species vs. bill_depth_mm? What
   might be a better choice of geom?

```{r}
ggplot(
  data = penguins,
  mapping = aes(
    x = species,
    y = bill_depth_mm,
    color = species,
    shape = species
  )
) +
  geom_point(na.rm = TRUE) +
  labs(
    title = "Penguine Bill Depth by Species",
    x = "Species",
    y = "Bill Depth (mm)",
    color = "Species", shape = "Species"
  )

ggplot(
  data = penguins,
  mapping = aes(x = species, y = bill_depth_mm, color = species)
) +
  geom_boxplot() +
  labs(
    title = "Penguine Bill Depth by Species",
    x = "Species",
    y = "Bill Depth (mm)",
    color = "Species"
  )
```

Obligatory link to https://nightingaledvs.com/ive-stopped-using-box-plots-should-you/

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = species, y = bill_depth_mm, color = species)
) +
  geom_jitter(alpha = 0.3, height = 0, width = 0.1, na.rm = TRUE) +
  labs(
    title = "Penguine Bill Depth by Species",
    x = "Species",
    y = "Bill Depth (mm)",
    color = "Species"
  )
```

8. Recreate the following visualization. What aesthetic should bill_depth_mm be
   mapped to? And should it be mapped at the global level or at the geom level?

```{r}
ggplot(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = bill_depth_mm), na.rm = TRUE) +
  geom_smooth(na.rm = TRUE)
```
