# Layers

## Facets

```{r}
#| output: false

library(tidyverse)
library(palmerpenguins)
```

### Exercises

> 1. What happens if you facet on a continuous variable?

I don't know, but the first chapter told us they have to be categorical! My
guess is an error.

```{r}
penguins |>
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(vars(flipper_length_mm))
```

No, it just treats every value as its own category. Very cute.

> 2. What do the empty cells in the plot above with `facet_grid(drv ~ cyl)`
>    mean? Run the following code. How do they relate to the resulting plot?

Well, they represent combinations of `drv` and `cyl` that do not appear in the
given data set.

The plot referred to in the question:

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  facet_grid(drv ~ cyl, scales = "free")
```

And the "following code":

```{r}
ggplot(mpg) +
  geom_point(aes(x = drv, y = cyl))
```

This demonstrates the locations at which there would be empty facets. No 7
cylinder vehicles, and no 4 cylinder RWDs, or 5 cylinder RWDs or 4WDs.

> 3. What plots does the following code make? What does `.` do?

Well, as per the docs, it seems that sending a formula to the `facet_*`
functions is a compatibility mode, and we should be setting `rows` and
`columns`, which is what I will do. I am still learning about formulae in R,
but I believe that a `.` means that there is nothing on that side of the
formula, and that there should be no facets there. I would hence expect the
first plot to only facet along rows, and the second to only facet along columns.

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

> 4. Take the first faceted plot in this section:

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(~ cyl, nrow = 2)
```

> What are the advantages to using faceting instead of the color aesthetic?
> What are the disadvantages? How might the balance change if you had a larger
> dataset?

Faceting and colour both attempt to visualise a third dimension in an otherwise
two-dimensional plot. Colours allow us to see clustering within a dataset,
whereas facets allow us to compare trends in places where colours might
overlap. Each facet also shows less data per plot, which will help keep plots
cleaner as a data set grows.

> 5. Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other
>    options control the layout of the individual panels? Why doesn’t
>    `facet_grid()` have `nrow` and `ncol` arguments?

`facet_wrap()` only facets on a single dimension, so it can be given a specific
number of rows and columns to wrap to. `facet_grid()` gets no such affordance
since its dimensions are defined by the size of the facets given to each axis.

`dir` also controls the layout of the individual panels, and I guess the scale
and axis arguments do as well. I don't know man, it's a vague question.

> 6. Which of the following plots makes it easier to compare engine size
>    (`displ`) across cars with different drive trains? What does this say
>    about when to place a faceting variable across rows or columns?

Just looking at the code, I'm going to guess that having multiple rows so that
our histograms align vertically is going to make this easier to look at.

```{r}
ggplot(mpg, aes(x = displ)) +
  geom_histogram() +
  facet_grid(drv ~ .)

ggplot(mpg, aes(x = displ)) +
  geom_histogram() +
  facet_grid(. ~ drv)
```

Oh boy was I right.

> 7. Recreate the following plot using `facet_wrap()` instead of
>    `facet_grid().` How do the positions of the facet labels change?

```{r}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)
```

Hmm.

```{r}
ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy)) +
  facet_wrap(vars(drv), dir = "v", strip.position = "right")
```

Well, the facet labels were up top by default.
